On the power of conditional independence testing under model-X \support{EK was partially supported by NSF DMS-2113072.}

### 2022-12-19 test/crtpower-ejs.tex
Added files for journal submission

\begin{theorem} \label{prop:crt-optimality}
	Let $\bar{\mathcal L} \in \mathscr L^{\text{MX}}(f^*)$ be an alternative distribution, with $\bar{\mathcal L}(\pry|\prx,\prz) = \bar f_{\pry|\prx,\prz}$. The likelihood of the data $(\srx, \sry, \srz)$ with respect to $\bar{\mathcal L}(\pry|\prx,\prz)$ is
	\begin{equation}
	T^{\textnormal{opt}}(\srx, \sry, \srz) \equiv \prod_{i = 1}^n  \bar f(\sry_i|\srx_i, \srz_i).
	\label{log-likelihood-ratio}
	\end{equation}
	The CRT $\phi^{\CRT}_{T^{\textnormal{opt}}}$ based on this test statistic is the most powerful conditionally valid test of $H_0: \mathcal L \in \mathscr L_0^{\textnormal{MX}}(f^*)$ against $H_1: \mathcal L = \bar{\mathcal L}$, i.e. 
	\begin{equation}
	\mathbb E_{\bar{\mathcal L}}[\phi(\srx,\sry,\srz)] \leq \mathbb E_{\bar{\mathcal L}}[\phi^{\CRT}_{T^\textnormal{opt}}(\srx, \sry, \srz)]
	\label{unconditional}
	\end{equation}
	for any test $\phi$ satisfying the conditional validity property~\eqref{eq:conditional-validity}.
\end{theorem}

@@

Now, consider a set of semiparametric (partially linear) alternatives $\mathcal L_n(\pry|\prx,\prz)$ such that
\begin{equation}
	\pry = \prx^T \beta_n + g_n(\prz) + \peps; \quad \peps \sim N(0, \sigma^2 ),\ \sigma^2 > 0
	\label{eq:linearity}
\end{equation}
for $\peps \independent (\prx, \pry, \prz)$. Here, $\beta_n \in \mathbb R^d$ is a coefficient vector, $g_n: \mathbb R^p \rightarrow \mathbb R$ a general function, and $\sigma^2 > 0$ the residual variance. Of special interest are local alternatives where $\beta_n = h/\sqrt{n}$ for some $h \in \mathbb R^d$. We emphasize that---in this section and throughout the paper---we use the partially linear model~\eqref{eq:linearity} exclusively as an alternative distribution against which to assess power, rather than an additional assumption required for Type-I error control. By Theorem~\ref{prop:crt-optimality}, the most powerful test against the alternative~\eqref{eq:linearity} is the CRT based on the likelihood statistic 
\begin{equation}
\begin{split}
T_n^{\text{opt}}(\srx, \sry, \srz) &= \prod_{i = 1}^n\frac{1}{(2\pi)^{1/2}}\exp\left(-\frac{1}{2\sigma^2}\left(\sry_i - \srx_i^T \beta_n - g_n(\srz_i)\right)^2\right) \\
&=\prod_{i = 1}^n\frac{1}{(2\pi)^{1/2}}\exp\left(-\frac{1}{2\sigma^2}\left(\sry_i - (\srx_i-\mu_n(\srz_i))^T \beta_n - g_n'(\srz_i)\right)^2\right),
\label{eq:likelihood-ratio}
\end{split}
\end{equation}
where
\begin{equation}
\bar g_n(\prz) \equiv \mathbb E[\pry|\prz] =  \mu_n(\prz)^T\beta_n  + g_n(\prz).
\label{eq:g-n-prime-def}
\end{equation}
Assuming local alternatives $\beta_n = h/\sqrt n$ and taking a logarithm, we obtain
\begin{equation}
	\begin{split}
		\log T_n^{\text{opt}}(\srx, \sry, \srz) &= -\frac n 2 \log(2\pi) - \frac{1}{2\sigma^2}\sum_{i = 1}^n\left(\sry_i - (\srx_i-\mu_n(\srz_i))^T h/\sqrt{n} - \bar g_n(\srz_i)\right)^2 \\
		&\approx   \frac{h^T}{\sigma^2}\frac{1}{\sqrt{n}}\sum_{i = 1}^n(\sry_i - \bar g_n(\srz_i))(\srx_i-\mu_n(\srz_i)) + C,
	\end{split}
\label{eq:optimal-semiparametric}
\end{equation}
where $C$ is a constant that does not depend on $\srx$ and therefore does not change upon resampling.

@@

If the conditional distribution $\mathcal L_n(\prx|\prz)$ is not Gaussian, then the dCRT $\phi_n^{\mathcal L_n}$ based on $T_n(\srx, \sry, \srz)$ will not reduce to the closed-form expression~\eqref{eq:crt-closed-form}. However, we can think of the test $\phi^{N(\mu_n, \Sigma_n)}_n$ as a kind of second-order approximation for $\phi_n^{\mathcal L_n}$ as long as $\mathcal L_n(\prx|\prz)$ has first and second moments given by $\mu_n(\prz)$ and $\Sigma_n(\prz)$, respectively. Indeed, it is easy to check that the resampling distribution $\mathcal L_n(\sqrt n \cdot \widehat \rho_n(\srxk, \sry, \srz) \mid \srx, \sry, \srz)$ matches that derived in the normal case~\eqref{eq:normal-resampling-distribution} up to two moments. Under a few assumptions, we can make this intuition precise by showing that $\phi_n^{\mathcal L_n}$ is asymptotically equivalent to $\phi^{N(\mu_n, \Sigma_n)}_n$ (Theorem~\ref{thm:equivalence} below). We require the distribution $\mathcal L_n$ to satisfy the following moment conditions \footnote{The exponents in these moment conditions can be relaxed from 4 to $2+\delta$, in particular, requiring an appropriate triangular array weak law of large numbers with 1+$\delta$ moments. This slight weakening of moment conditions requires significantly more technical effort, so is omitted for simplicity since it does not alter the main takeaway messages of our analysis.}for fixed $c_1, c_2 > 0$:
\small
\begin{equation}
\mathcal L_n \in \mathscr L_n(c_1, c_2) \equiv \{\mathcal L_n: \|S_{n}^{-1}\| \leq c_1, \mathbb E_{\mathcal L_n}\left[(\pry - \widehat g_n(\prz))^{4} \mathbb E_{\mathcal L_n}[\|\prx - \mu_n(\prz)\|^{4}|\prz]\right] \leq c_2 \},
\label{eq:moment-conditions}
\end{equation} 
\normalsize
where
\begin{equation}
S_n^2 \equiv \mathbb E[\widehat S_n^2] = \mathbb E_{\mathcal L_n}\left[(\pry - \widehat g_n(\prz))^2 \Sigma_n(\prz)\right].
\end{equation}
Furthermore, to avoid technical complications, we assume that the estimate $\widehat g_n$ is trained on an independent dataset (whose size can vary arbitrarily with $n$ and is not included in the sample size $n$ used for testing). For example, there has been recent interest in combining observational and experimental (randomized) data; typically, the former is much more abundant than the latter. We can think of $\widehat g_n$ being trained on the former, and then used for MX inference on the latter \citep{Bates2020}.
These training sets across $n$ and resulting estimates $\widehat g_n$ remain fixed throughout.

@@

The MX(2) $F$-test controls the Type-I error under the MX(2) assumption, if the moment conditions~\eqref{eq:moment-conditions} hold and $\widehat g_n$ is fit out of sample.
\begin{theorem} \label{thm:asymptotic-alpha-level}
If $\mathcal L_n \in \mathscr L_0^{\textnormal{MX(2)}}\cap \mathscr L_n(c_1, c_2)$ for some $c_1, c_2 > 0$ and $\widehat g_n$ is fit out of sample, then the standardized quantity $U_n(\srx, \sry, \srz)$ converges to the standard normal: 
\begin{equation}
	U_n(\srx, \sry, \srz) \overset{\mathcal L_n}\rightarrow_d N(0, I_d).
	\label{eq:asymptotic-normality}
\end{equation}
Therefore, the MX(2) $F$-test controls Type-I error asympotically, uniformly over the above subset of $\mathscr L_0^{\textnormal{MX(2)}}$: 
\begin{equation}
	\limsup_{n \rightarrow \infty} \sup_{\mathcal L_n \in \mathscr L_0^{\textnormal{MX(2)}}\cap \mathscr L_n(c_1, c_2)} \mathbb E_{\mathcal L_n}[\phi^{N(\mu_n, \Sigma_n)}_n(\srx, \sry, \srz)] \leq \alpha.
	\label{asymptotic-alpha-level}
\end{equation}
\end{theorem}

@@

Finally, we note a connection between Theorem~\ref{thm:power} and causal inference. It is widely known in causal inference (see e.g. \citet[Section 7.5]{Imbens2015}) that adjustment for covariates $\srz$ in randomized experiments (a) yields consistent estimates despite misspecification of $\mathcal L(\pry|\prx,\prz)$ and (b) improve estimation efficiency to the extent that this adjustment captures the distribution $\mathcal L(\pry|\prx,\prz)$. This fact mirrors the conclusions of Theorem~\ref{thm:power}. The asymptotic variance of the regression-based estimator for the average treatment effect in a completely randomized experiment is a standard result, but we are unaware of a quantitative expression of the asympotic efficiency of covariate-adjusted versions of the Fisher randomization test (though some insight is provided by \cite{Zhao2021}).

@@

\begin{theorem} \label{prop:knockoff-optimality}
	Let $\bar{\mathcal L}$ be a fixed alternative distribution for $(\prx,\pry)$, with $\bar{\mathcal L}(\pry|\prx) = \bar f(\pry|\prx)$. Define the likelihood statistic
	\begin{equation}
	T_j^{\textnormal{opt}}([\srx, \srxk], \sry) \equiv \prod_{i = 1}^n\bar f(\sry_i|\srx_{i,\bullet}).
	\label{log-likelihood-ratio-knockoffs}
	\end{equation}
	Assuming that ties do not occur, that is
	\begin{equation}
	\mathbb P_{\bar {\mathcal L}}[T^{\textnormal{opt}}_j([\srx, \srxk], \sry) = T^{\textnormal{opt}}_j([\srx, \srxk]_{\textnormal{swap}(j)}, \sry), \srx_{\bullet, j} \neq \srxk_{\bullet, j}] = 0, 
	\label{nondegeneracy-assumption}
	\end{equation}
	we have that the above likelihood statistic yields an optimal one-bit $p$-value:
	\begin{equation}
	T_j^{\textnormal{opt}} \in \underset{T_j}{\arg \max}\ \mathbb P[T_j([\srx, \srxk], \sry) > T_j([\srx, \srxk]_{\textnormal{swap}(j)}, \sry)].
	\label{unconditional-knockoff-optimality}
	\end{equation}
\end{theorem}

@@

\bibliographystyle{imsart-nameyear}
\bibliography{/Users/ekatsevi/.references/crtpower}

@@

	Next, note that for any $\delta > 0$,
	\small
	\begin{equation*}
		\begin{split}
			&\mathbb P_{\mathcal L_n}[\phi^{N(\mu_n, \Sigma_n)}_n(\srx, \sry, \srz) \neq \phi^{\mathcal L_n}_n(\srx, \sry, \srz)] \\
			&\quad = \mathbb P_{\mathcal L_n}[\min(c_{d,1-\alpha},C_n(\sry, \srz))  < T_n(\srx, \sry, \srz) \leq \max(c_{d,1-\alpha},C_n(\sry, \srz))] \\
			&\quad=\mathbb P_{\mathcal L_n}[\min(c_{d,1-\alpha},C_n(\sry, \srz))  < T_n(\srx, \sry, \srz) \leq \max(c_{d,1-\alpha},C_n(\sry, \srz)), |C_n(\sry, \srz)-c_{d,1-\alpha}| \leq \delta] \\
			&\quad \quad +  \mathbb P_{\mathcal L_n}[\min(c_{d,1-\alpha},C_n(\sry, \srz))  < T_n(\srx, \sry, \srz) \leq \max(c_{d,1-\alpha},C_n(\sry, \srz)), |C_n(\sry, \srz)-c_{d,1-\alpha}| > \delta]\\
			&\quad\leq \mathbb P_{\mathcal L_n}[|T_n(\srx, \sry, \srz)-c_{d,1-\alpha}| \leq \delta] + \mathbb P_{\mathcal L_n}[|C_n(\sry, \srz)-c_{d,1-\alpha}| > \delta].
		\end{split}
	\end{equation*}
	\normalsize
	To justify the last step, suppose without loss of generality that $c_{d,1-\alpha} \leq C_n(\sry, \srz)$. Then, note that if $c_{d,1-\alpha} < T_n(\srx, \sry, \srz) \leq C_n(\sry, \srz)$ and $C_n(\sry, \srz)-c_{d,1-\alpha} \leq \delta$ then
	\begin{equation*}
		|T_n(\srx, \sry,\srz)-c_{d,1-\alpha}| = T_n(\srx, \sry,\srz)-c_{d,1-\alpha} \leq C_n(\sry, \srz)- c_{d,1-\alpha} \leq \delta.
	\end{equation*}
	Taking a $\limsup$ on both sides in the display before the last and using the convergence $C_n(\sry,\srz)  \overset{\mathcal L_n}\rightarrow_p c_{d,1-\alpha}$, we find that
	\begin{equation*}
		\begin{split}
			&\limsup_{n \rightarrow \infty}\ \mathbb P_{\mathcal L_n}[\phi^{N(\mu_n, \Sigma_n)}_n(\srx, \sry, \srz) \neq \phi^{\mathcal L_n}_n(\srx, \sry, \srz)] \\
			&\quad \leq \limsup_{n \rightarrow \infty}\ \mathbb P_{\mathcal L_n}[|T_n(\srx, \sry, \srz)-c_{d,1-\alpha}| \leq \delta]+ \limsup_{n \rightarrow \infty}\ \mathbb P_{\mathcal L_n}[|C_n(\sry, \srz)-c_{d,1-\alpha}| > \delta] \\
			&\quad = \limsup_{n \rightarrow \infty}\ \mathbb P_{\mathcal L_n}[|T_n(\srx, \sry, \srz)-c_{d,1-\alpha}| \leq \delta].
		\end{split}
	\end{equation*}
	Letting $\delta \rightarrow 0$ and using the assumption~\eqref{eq:non-accumulation}, we arrive at the claimed asymptotic equivalence~\eqref{eq:asymptotic-equivalence}. This completes the proof.
\end{proof}

@@

To show the asymptotic Type-I error control statement~\eqref{asymptotic-alpha-level}, it suffices to show that for any sequence $\mathcal L_n \in \mathscr L_0^{\textnormal{MX(2)}}\cap \mathscr L_n(c_1, c_2)$, we have
	\begin{equation}
		\limsup_{n \rightarrow \infty}\ \mathbb E_{\mathcal L_n}[\phi^{N(\mu_n, \Sigma_n)}_n(\srx, \sry, \srz)] \leq \alpha.
	\label{eq:pointwise-alpha-level}
	\end{equation}
By the continuous mapping theorem it follows from asymptotic normality~\eqref{eq:asymptotic-normality} that $T_n(\srx,\sry,\srz) = \|U_n(\srx, \sry,\srz)\|^2 \overset{\mathcal L_n}\rightarrow_d \chi^2_d$. Therefore,
	\begin{equation*}
		\lim_{n \rightarrow \infty}\ \mathbb E_{\mathcal L_n}[\phi^{N(\mu_n, \Sigma_n)}_n(\srx, \sry, \srz)] = \lim_{n \rightarrow \infty}\ \mathbb P_{\mathcal L_n}[T_n(\srx,\sry,\srz) > c_{d,1-\alpha}] = \mathbb P[\chi^2_d > c_{d,1-\alpha}] = \alpha,
	\end{equation*}
	from which the conclusion~\eqref{eq:pointwise-alpha-level} follows. This completes the proof.
\end{proof}

@@

First, we state a lemma that gives a sufficient condition for the convergence of the CRT threshold, which follows directly from  Lemmas 2 and 3 of \cite{Wang2020b}.
\begin{lemma}[\cite{Wang2020b}] \label{lem:lucas}
	Let $\mathcal L_n$ be a sequence of laws over $(\prx,\pry,\prz)$, from which $(\srx,\sry,\srz)$ are sampled. Furthermore, for each $i$ sample two independent copies
	\begin{equation}
		\srxk_i^1,\srxk_i^2 \overset{\text{i.i.d.}}\sim \mathcal L_n(\prx|\prz = \srz_i) \quad \text{such that, given } \srz,  (\srxk_1^1,\srxk_1^2) \independent \cdots \independent (\srxk_n^1,\srxk_n^2) \independent \sry.
		\label{eq:two-resamples}
	\end{equation}
	Suppose that $T_n(\srx,\sry,\srz)$ is a test statistic satisfying 
	\begin{equation}
		(T_n(\srxk^1,\sry,\srz), T_n(\srxk^2, \sry,\srz)) \overset{\mathcal L_n}\rightarrow_d  \widetilde T \times \widetilde T
	\end{equation}
	for some limiting random variable $\widetilde T$ with continuous and strictly increasing distribution function. Then, the CRT threshold converges in probability to the upper quantile of $\widetilde T$:
	\begin{equation}
		C_n(\sry, \srz) \equiv Q_{1-\alpha}[T_n(\srxk,\sry,\srz)|\sry,\srz]  \overset{\mathcal L_n}\rightarrow_p Q_{1-\alpha}[\widetilde T].
	\end{equation}

@@

			&\quad\leq C\mathbb E_{\mathcal L_n}\left[(((\prx - \mu_n(\prz))^T \beta_n)^4 + \peps^4 + (\bar g_n(\prz)-\widehat g_n(\prz))^4) \mathbb E_{\mathcal L_n}[\|\prx - \mu_n(\prz)\|^{4}|\prz]\right] \\

>>>>

\begin{theorem} \label{prop:crt-optimality}
	Let $\bar{\mathcal L} \in \mathscr L^{\text{MX}}(f^*)$ be an alternative distribution, with conditional distribution $\bar{\mathcal L}(\pry|\prx,\prz) = \bar f_{\pry|\prx,\prz}$. The likelihood of the data $(\srx, \sry, \srz)$ with respect to $\bar{\mathcal L}(\pry|\prx,\prz)$ is
	\begin{equation}
	T^{\textnormal{opt}}(\srx, \sry, \srz) \equiv \prod_{i = 1}^n  \bar f(\sry_i|\srx_i, \srz_i).
	\label{log-likelihood-ratio}
	\end{equation}
	The CRT $\phi^{\CRT}_{T^{\textnormal{opt}}}$ based on this test statistic is the most powerful conditionally valid test of $H_0: \mathcal L \in \mathscr L_0^{\textnormal{MX}}(f^*)$ against $H_1: \mathcal L = \bar{\mathcal L}$, i.e. 
	\begin{equation}
	\mathbb E_{\bar{\mathcal L}}[\phi(\srx,\sry,\srz)] \leq \mathbb E_{\bar{\mathcal L}}[\phi^{\CRT}_{T^\textnormal{opt}}(\srx, \sry, \srz)]
	\label{unconditional}
	\end{equation}
	for any test $\phi$ satisfying the conditional validity property~\eqref{eq:conditional-validity}.
\end{theorem}

@@

Now, consider a set of semiparametric (partially linear) alternative distributions $\mathcal L_n(\pry|\prx,\prz)$ such that
\begin{equation}
	\pry = \prx^T \beta_n + g_n(\prz) + \peps; \quad \peps \sim N(0, \sigma^2 ),\ \sigma^2 > 0
	\label{eq:linearity}
\end{equation}
for $\peps \independent (\prx, \pry, \prz)$. Here, $\beta_n \in \mathbb R^d$ is a coefficient vector, $g_n: \mathbb R^p \rightarrow \mathbb R$ a general function, and $\sigma^2 > 0$ the residual variance. Of special interest are local alternatives where $\beta_n = h/\sqrt{n}$ for some $h \in \mathbb R^d$. We emphasize that---in this section and throughout the paper---we use the partially linear model~\eqref{eq:linearity} exclusively as an alternative distribution against which to assess power, rather than an additional assumption required for Type-I error control. By Theorem~\ref{prop:crt-optimality}, the most powerful test against the alternative~\eqref{eq:linearity} is the CRT based on the likelihood statistic 
\begin{equation}
\begin{split}
T_n^{\text{opt}}(\srx, \sry, \srz) &= \prod_{i = 1}^n\frac{1}{(2\pi)^{1/2}}\exp\left(-\frac{1}{2\sigma^2}\left(\sry_i - \srx_i^T \beta_n - g_n(\srz_i)\right)^2\right) \\
&=\prod_{i = 1}^n\frac{1}{(2\pi)^{1/2}}\exp\left(-\frac{1}{2\sigma^2}\left(\sry_i - (\srx_i-\mu_n(\srz_i))^T \beta_n - g_n'(\srz_i)\right)^2\right),
\label{eq:likelihood-ratio}
\end{split}
\end{equation}
where
\begin{equation}
\bar g_n(\prz) \equiv \mathbb E[\pry|\prz] =  \mu_n(\prz)^T\beta_n  + g_n(\prz).
\label{eq:g-n-prime-def}
\end{equation}
Assuming local alternatives $\beta_n = h/\sqrt n$ and taking a logarithm, we obtain
\begin{equation}
	\begin{split}
		&\log T_n^{\text{opt}}(\srx, \sry, \srz) \\
		&\quad= -\frac n 2 \log(2\pi) - \frac{1}{2\sigma^2}\sum_{i = 1}^n\left(\sry_i - (\srx_i-\mu_n(\srz_i))^T h/\sqrt{n} - \bar g_n(\srz_i)\right)^2 \\
		&\quad\approx   \frac{h^T}{\sigma^2}\frac{1}{\sqrt{n}}\sum_{i = 1}^n(\sry_i - \bar g_n(\srz_i))(\srx_i-\mu_n(\srz_i)) + C,
	\end{split}
\label{eq:optimal-semiparametric}
\end{equation}
where $C$ is a constant that does not depend on $\srx$ and therefore does not change upon resampling.

@@

If the conditional distribution $\mathcal L_n(\prx|\prz)$ is not Gaussian, then the dCRT $\phi_n^{\mathcal L_n}$ based on $T_n(\srx, \sry, \srz)$ will not reduce to the closed-form expression~\eqref{eq:crt-closed-form}. However, we can think of the test $\phi^{N(\mu_n, \Sigma_n)}_n$ as a kind of second-order approximation for $\phi_n^{\mathcal L_n}$ as long as $\mathcal L_n(\prx|\prz)$ has first and second moments given by $\mu_n(\prz)$ and $\Sigma_n(\prz)$, respectively. Indeed, it is easy to check that the resampling distribution $\mathcal L_n(\sqrt n \cdot \widehat \rho_n(\srxk, \sry, \srz) \mid \srx, \sry, \srz)$ matches that derived in the normal case~\eqref{eq:normal-resampling-distribution} up to two moments. Under a few assumptions, we can make this intuition precise by showing that $\phi_n^{\mathcal L_n}$ is asymptotically equivalent to $\phi^{N(\mu_n, \Sigma_n)}_n$ (Theorem~\ref{thm:equivalence} below). We require the distribution $\mathcal L_n$ to satisfy the following moment conditions \footnote{The exponents in these moment conditions can be relaxed from 4 to $2+\delta$, in particular, requiring an appropriate triangular array weak law of large numbers with 1+$\delta$ moments. This slight weakening of moment conditions requires significantly more technical effort, so is omitted for simplicity since it does not alter the main takeaway messages of our analysis.}for fixed $c_1, c_2 > 0$:
\small
\begin{equation}
\begin{split}
\mathcal L_n &\in \mathscr L_n(c_1, c_2) \\
&\equiv \{\mathcal L_n: \|S_{n}^{-1}\| \leq c_1, \mathbb E_{\mathcal L_n}\left[(\pry - \widehat g_n(\prz))^{4} \mathbb E_{\mathcal L_n}[\|\prx - \mu_n(\prz)\|^{4}|\prz]\right] \leq c_2 \},
\label{eq:moment-conditions}
\end{split}
\end{equation} 
\normalsize
where
\begin{equation}
S_n^2 \equiv \mathbb E[\widehat S_n^2] = \mathbb E_{\mathcal L_n}\left[(\pry - \widehat g_n(\prz))^2 \Sigma_n(\prz)\right].
\end{equation}
Furthermore, to avoid technical complications, we assume that the estimate $\widehat g_n$ is trained on an independent dataset (whose size can vary arbitrarily with $n$ and is not included in the sample size $n$ used for testing). For example, there has been recent interest in combining observational and experimental (randomized) data; typically, the former is much more abundant than the latter. We can think of $\widehat g_n$ being trained on the former, and then used for MX inference on the latter \citep{Bates2020}.
These training sets across $n$ and resulting estimates $\widehat g_n$ remain fixed throughout.

@@

The MX(2) $F$-test controls the Type-I error under the MX(2) assumption, if the moment conditions~\eqref{eq:moment-conditions} hold and $\widehat g_n$ is fit out of sample.
\begin{theorem} \label{thm:asymptotic-alpha-level}
If $\mathcal L_n \in \mathscr L_0^{\textnormal{MX(2)}}\cap \mathscr L_n(c_1, c_2)$ for some $c_1, c_2 > 0$ and $\widehat g_n$ is fit out of sample, then the standardized quantity $U_n(\srx, \sry, \srz)$ converges to the standard normal: 
\begin{equation}
	U_n(\srx, \sry, \srz) \overset{\mathcal L_n}\rightarrow_d N(0, I_d).
	\label{eq:asymptotic-normality}
\end{equation}
Therefore, the MX(2) $F$-test controls Type-I error asymptotically, uniformly over the above subset of $\mathscr L_0^{\textnormal{MX(2)}}$: 
\begin{equation}
	\limsup_{n \rightarrow \infty} \sup_{\mathcal L_n \in \mathscr L_0^{\textnormal{MX(2)}}\cap \mathscr L_n(c_1, c_2)} \mathbb E_{\mathcal L_n}[\phi^{N(\mu_n, \Sigma_n)}_n(\srx, \sry, \srz)] \leq \alpha.
	\label{asymptotic-alpha-level}
\end{equation}
\end{theorem}

@@

Finally, we note a connection between Theorem~\ref{thm:power} and causal inference. It is widely known in causal inference (see e.g. \citet[Section 7.5]{Imbens2015}) that adjustment for covariates $\srz$ in randomized experiments (a) yields consistent estimates despite misspecification of $\mathcal L(\pry|\prx,\prz)$ and (b) improve estimation efficiency to the extent that this adjustment captures the distribution $\mathcal L(\pry|\prx,\prz)$. This fact mirrors the conclusions of Theorem~\ref{thm:power}. The asymptotic variance of the regression-based estimator for the average treatment effect in a completely randomized experiment is a standard result, but we are unaware of a quantitative expression of the asymptotic efficiency of covariate-adjusted versions of the Fisher randomization test (though some insight is provided by \cite{Zhao2021}).

@@

\begin{theorem} \label{prop:knockoff-optimality}
	Let $\bar{\mathcal L}$ be a fixed alternative distribution for $(\prx,\pry)$, with conditional $\bar{\mathcal L}(\pry|\prx) = \bar f(\pry|\prx)$. Define the likelihood statistic
	\begin{equation}
	T_j^{\textnormal{opt}}([\srx, \srxk], \sry) \equiv \prod_{i = 1}^n\bar f(\sry_i|\srx_{i,\bullet}).
	\label{log-likelihood-ratio-knockoffs}
	\end{equation}
	Assuming that ties do not occur, that is
	\begin{equation}
	\mathbb P_{\bar {\mathcal L}}[T^{\textnormal{opt}}_j([\srx, \srxk], \sry) = T^{\textnormal{opt}}_j([\srx, \srxk]_{\textnormal{swap}(j)}, \sry), \srx_{\bullet, j} \neq \srxk_{\bullet, j}] = 0, 
	\label{nondegeneracy-assumption}
	\end{equation}
	we have that the above likelihood statistic yields an optimal one-bit $p$-value:
	\begin{equation}
	T_j^{\textnormal{opt}} \in \underset{T_j}{\arg \max}\ \mathbb P[T_j([\srx, \srxk], \sry) > T_j([\srx, \srxk]_{\textnormal{swap}(j)}, \sry)].
	\label{unconditional-knockoff-optimality}
	\end{equation}
\end{theorem}

@@

\bibliographystyle{imsart-nameyear}
\bibliography{crtpower}

@@

	Next, note that for any $\delta > 0$,
	\small
	\begin{align*}
			&\mathbb P_{\mathcal L_n}[\phi^{N(\mu_n, \Sigma_n)}_n(\srx, \sry, \srz) \neq \phi^{\mathcal L_n}_n(\srx, \sry, \srz)] \\
			&\quad = \mathbb P_{\mathcal L_n}[\min(c_{d,1-\alpha},C_n(\sry, \srz))  < T_n(\srx, \sry, \srz) \leq \max(c_{d,1-\alpha},C_n(\sry, \srz))] \\
			&\quad=\mathbb P_{\mathcal L_n}[\min(c_{d,1-\alpha},C_n(\sry, \srz))  < T_n(\srx, \sry, \srz) \leq \max(c_{d,1-\alpha},C_n(\sry, \srz)), \\
			& \quad \quad \qquad \ |C_n(\sry, \srz)-c_{d,1-\alpha}| \leq \delta] \\
			&\quad \quad +  \mathbb P_{\mathcal L_n}[\min(c_{d,1-\alpha},C_n(\sry, \srz))  < T_n(\srx, \sry, \srz) \leq \max(c_{d,1-\alpha}, C_n(\sry, \srz)), \\
			& \quad \quad \quad \qquad \  |C_n(\sry, \srz)-c_{d,1-\alpha}| > \delta]\\
			&\quad\leq \mathbb P_{\mathcal L_n}[|T_n(\srx, \sry, \srz)-c_{d,1-\alpha}| \leq \delta] + \mathbb P_{\mathcal L_n}[|C_n(\sry, \srz)-c_{d,1-\alpha}| > \delta].
	\end{align*}
	\normalsize
	To justify the last step, suppose without loss of generality that $c_{d,1-\alpha} \leq C_n(\sry, \srz)$. Then, note that if $c_{d,1-\alpha} < T_n(\srx, \sry, \srz) \leq C_n(\sry, \srz)$ and $C_n(\sry, \srz)-c_{d,1-\alpha} \leq \delta$ then
	\begin{equation*}
		|T_n(\srx, \sry,\srz)-c_{d,1-\alpha}| = T_n(\srx, \sry,\srz)-c_{d,1-\alpha} \leq C_n(\sry, \srz)- c_{d,1-\alpha} \leq \delta.
	\end{equation*}
	Taking a $\limsup$ on both sides in the display before the last and using the convergence $C_n(\sry,\srz)  \overset{\mathcal L_n}\rightarrow_p c_{d,1-\alpha}$, we find that
	\begin{equation*}
		\begin{split}
			&\limsup_{n \rightarrow \infty}\ \mathbb P_{\mathcal L_n}[\phi^{N(\mu_n, \Sigma_n)}_n(\srx, \sry, \srz) \neq \phi^{\mathcal L_n}_n(\srx, \sry, \srz)] \\
			&\quad \leq \limsup_{n \rightarrow \infty}\ \mathbb P_{\mathcal L_n}[|T_n(\srx, \sry, \srz)-c_{d,1-\alpha}| \leq \delta] \\
			& \quad \quad + \limsup_{n \rightarrow \infty}\ \mathbb P_{\mathcal L_n}[|C_n(\sry, \srz)-c_{d,1-\alpha}| > \delta] \\
			&\quad = \limsup_{n \rightarrow \infty}\ \mathbb P_{\mathcal L_n}[|T_n(\srx, \sry, \srz)-c_{d,1-\alpha}| \leq \delta].
		\end{split}
	\end{equation*}
	Letting $\delta \rightarrow 0$ and using the assumption~\eqref{eq:non-accumulation}, we arrive at the claimed asymptotic equivalence~\eqref{eq:asymptotic-equivalence}. This completes the proof.
\end{proof}

@@

To show the asymptotic Type-I error control statement~\eqref{asymptotic-alpha-level}, it suffices to show that for any sequence $\mathcal L_n \in \mathscr L_0^{\textnormal{MX(2)}}\cap \mathscr L_n(c_1, c_2)$, we have
	\begin{equation}
		\limsup_{n \rightarrow \infty}\ \mathbb E_{\mathcal L_n}[\phi^{N(\mu_n, \Sigma_n)}_n(\srx, \sry, \srz)] \leq \alpha.
	\label{eq:pointwise-alpha-level}
	\end{equation}
By the continuous mapping theorem it follows from asymptotic normality~\eqref{eq:asymptotic-normality} that $T_n(\srx,\sry,\srz) = \|U_n(\srx, \sry,\srz)\|^2 \overset{\mathcal L_n}\rightarrow_d \chi^2_d$. Therefore,
	\begin{align*}
		\lim_{n \rightarrow \infty}\ \mathbb E_{\mathcal L_n}[\phi^{N(\mu_n, \Sigma_n)}_n(\srx, \sry, \srz)] &= \lim_{n \rightarrow \infty}\ \mathbb P_{\mathcal L_n}[T_n(\srx,\sry,\srz) > c_{d,1-\alpha}] \\
		&= \mathbb P[\chi^2_d > c_{d,1-\alpha}] = \alpha,
	\end{align*}
	from which the conclusion~\eqref{eq:pointwise-alpha-level} follows. This completes the proof.
\end{proof}

@@

First, we state a lemma that gives a sufficient condition for the convergence of the CRT threshold, which follows directly from  Lemmas 2 and 3 of \cite{Wang2020b}.
\begin{lemma}[\cite{Wang2020b}] \label{lem:lucas}
	Let $\mathcal L_n$ be a sequence of laws over $(\prx,\pry,\prz)$, from which $(\srx,\sry,\srz)$ are sampled. Furthermore, for each $i$ sample two independent copies
	\begin{equation}
		\srxk_i^1,\srxk_i^2 \overset{\text{i.i.d.}}\sim \mathcal L_n(\prx|\prz = \srz_i)
		\label{eq:two-resamples}
	\end{equation}
	such that, given $\srz$,
	\begin{equation*}
		(\srxk_1^1,\srxk_1^2) \independent \cdots \independent (\srxk_n^1,\srxk_n^2) \independent \sry.
	\end{equation*}
	Suppose that $T_n(\srx,\sry,\srz)$ is a test statistic satisfying 
	\begin{equation}
		(T_n(\srxk^1,\sry,\srz), T_n(\srxk^2, \sry,\srz)) \overset{\mathcal L_n}\rightarrow_d  \widetilde T \times \widetilde T
	\end{equation}
	for some limiting random variable $\widetilde T$ with continuous and strictly increasing distribution function. Then, the CRT threshold converges in probability to the upper quantile of $\widetilde T$:
	\begin{equation}
		C_n(\sry, \srz) \equiv Q_{1-\alpha}[T_n(\srxk,\sry,\srz)|\sry,\srz]  \overset{\mathcal L_n}\rightarrow_p Q_{1-\alpha}[\widetilde T].
	\end{equation}

@@

			&\quad\leq C\mathbb E_{\mathcal L_n}\left[(((\prx - \mu_n(\prz))^T \beta_n)^4\mathbb E_{\mathcal L_n}[\|\prx - \mu_n(\prz)\|^{4}|\prz]\right] \\
			&\quad \quad + C\mathbb E_{\mathcal L_n}\left[\peps^4 \mathbb E_{\mathcal L_n}[\|\prx - \mu_n(\prz)\|^{4}|\prz]\right] \\
			& \quad \quad + C\mathbb E_{\mathcal L_n}\left[(\bar g_n(\prz)-\widehat g_n(\prz))^4\mathbb E_{\mathcal L_n}[\|\prx - \mu_n(\prz)\|^{4}|\prz]\right] \\
